---
title: "Introduction to train/predict/score"
author: "Michel Lang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to train/predict/score}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(mlr3)
knitr::opts_knit$set(
  datatable.print.keys = FALSE,
  datatable.print.class = TRUE
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this introduction we simply fit a classification tree on the iris and determine the mean misclassification error.

## Objects

First, we need to generate retrieve the following `mlr3` objects from the task dictionary and the learner dictionary, respectively:

1. The classification task
    ```{r}
    task = mlr_tasks$get("iris")
    ```
2. A learner for the classification tree
    ```{r}
    learner = mlr_learners$get("classif.rpart")
    ```

## Splitting into train and test

We opt to learn on roughly $\frac{4}{5}$ of the observations and predict on the other $\frac{1}{5}$.
To do so we create two index vectors:
```{r}
train.set = sample(task$nrow, 4/5 * task$nrow)
test.set = setdiff(seq_len(task$nrow), train.set)
```

## Training

Here, we call the `train()` function which returns an `Experiment` object:
```{r}
e = train(task, learner, subset = train.set)
print(e)
```

The printer also shows a summary of the state of the experiment.
We provided a task, a learner and have stored the model, but do not (yet) have predictions nor a performance measure.
We can easily access the object `rpart` has returned, e.g. to plot it or print it:
```{r}
rpart.model = e$model
print(rpart.model)
```

## Predicting

We take the incomplete experiment from before and declare for which observations of the task we want predictions.
```{r}
e = predict(e, subset = test.set)
print(e)
```
The predictions can be retrieved as a simple `data.table`.
```{r}
head(e$predictions)
```

## Performance Assessment

We augment the experiment again, now the performance measure:
```{r}
measure = mlr_measures$get("mmce")
e = score(e, measures = measure)
print(e)
e$performance$mmce
```

The experiment is now "complete" which means we can access all of its methods.


## Alternative interface

Instead of calling the functions `train()`, `predict()` and `score()`, it is also possible to directly work with Experiments:
```{r}
Experiment$new(task = task, learner = learner)$train(train.set)$predict(test.set)$score("mmce")
```

