---
title: "Introduction to Tasks"
author: "Michel Lang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Tasks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(mlr3)
knitr::opts_knit$set(
  datatable.print.keys = FALSE,
  datatable.print.class = TRUE
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(123)
```

## Predefined tasks

`mlr3` ships with some popular machine learning toy tasks.
These are stored in a `mlr3::Dictionary`, a simple key-value store, named `mlr3::mlr_tasks`:

```{r}
mlr_tasks

# list keys
names(mlr_tasks)

# get a quick overview
as.data.frame(mlr_tasks)
```

Here, we will stick to the popular [iris data set](https://en.wikipedia.org/wiki/Iris_flower_data_set):

```{r}
task = mlr_tasks$get("iris")
print(task)
```

Several task properties and characteristics can be queried using the task's methods (see `mlr3::Task`).
Most of them should be self-explanatory.

```{r}
task$nrow
task$ncol
task$target_names
task$class_n
task$class_names
task$feature_names
task$formula
task$head(n = 3)
```

In `mlr3`, each row has a unique identifier (row name) which can be either `integer` or `character.`
These can be used to select specific rows.

```{r}
# iris uses integer row_ids
task$row_ids

# filter rows 1, 51, and 101
task$data(rows = c(1, 51, 101))

# filter rows 1, 51, and 101 and only select column "Species"
task$data(rows = c(1, 51, 101), cols = "Species")
```


## Task Creation

To create an own task from a `data.frame`, you must determine the task type to select the respective constructor:

* Classification Task: Target column is labels (stored as `character`/`factor`) with only few distinct values.
  <br>$\Rightarrow$ `TaskClassif`.
* Regression Task: Target column is numeric (stored as `integer`/`double`).
  <br>$\Rightarrow$ `TaskRegr`.
* Cluster Task: You don't have a target but want to identify similarities in the feature space.
  <br>$\Rightarrow$ Not yet implemented.

Let's assume we want to create a simple regression task using the `mtcars` data set from the package `datasets` to predict the column `"mpg"` (miles per gallon).
We only take the first two features here to keep the output in the following examples short.

```{r}
data("mtcars", package = "datasets")
cars = mtcars[, 1:3]
str(cars)
```

Before we can create a regression task, we must create a `mlr3::DataBackend`, an abstraction for data storage system.
Here, we will stick to the simplest form of data storage: an in-memory table format using `data.table::data.table()`.
We construct the backend first, and then pass it to the regression task constructor:
```{r}
b = as_data_backend(cars)
task = TaskRegr$new(id = "cars", b, target = "mpg")
print(task)
```
The rownames of the `data.frame` `cars` have been automatically used as `row_ids`.
We can use analogously to integer row ids to filter observations of the task:
```{r}
task$row_ids
task$data(rows = c("Merc 280", "Volvo 142E"))
```


## Column Roles

Now, we want the original `rownames()` of `mtcars` to be a regular column.
Thus, we first pre-process the `data.frame` and then re-create the task.

```{r}
library("data.table")
# `as.data.table` removes rownames, but  argument `keep.rownames` ensures
# that they are stored in a separate column before
data = as.data.table(cars[, 1:3], keep.rownames = TRUE)
b = as_data_backend(data)
task = TaskRegr$new(id = "cars", b, target = "mpg")

# we now have integer row_ids
task$row_ids

# there is a new "feature" called "rn"
task$feature_names
```

In `mlr3`, columns (and rows) can be assigned roles.
We have seen three different roles for columns so far:

1. The target column (here `"mpg"`), also called dependent variable.
2. Features, also called independent variables.
3. The `row_id`. This column is there for technical reasons, and is typically useless for learning.

The different roles are stored as as a list of column name vectors:

```{r}
task$col_roles
```
As the output shows, the column is `"mpg"` is the target column and are three features: `"rn"` (rownames), `"cyl"`, and `"disp"`.
More roles are documented in the help for tasks.

In the following we do not want to learn on neither the primary key (which is taken care of `mlr3`) nor the new column `rn` with the row names.
However, we still might want to carry `rn` around for different reasons.
E.g., we can use the row names in plots or to associate outliers with the row names.
This being said, we need to change the role of the rownames column `rn` and remove it from the set of features.

```{r}
task$feature_names
task$set_col_role("rn", new_roles = character(0L))

# "rn" not listed as feature any more
task$feature_names

# also eliminated from "data" and "head"
task$data(rows = 1:2)
task$head(2)
```

## Row roles

Just like columns, you can also assign different roles to rows.
Rows can have two different roles:

1. Role `"use"`:
   Rows that are generally available for model fitting (although they may also be used as test set in resampling).
   This is the default role.
2. Role `"validation"`:
   Rows that are held back (see below).
   Rows which have missing values in the target column upon task creation are automatically moved to the validation set.

There are several reasons to hold some observations back or treat them differently:

1. It is often good practice to validate the final model on an external validation set to uncover possible overfitting
2. Some observations may be unlabeled, e.g. in data mining cups or [Kaggle](https://www.kaggle.com/) competitions.
   These observations cannot be used for training a model, but you can still predict labels.

Instead of creating a task with only a subset of observations and then manually apply the fitted model on an hold-back `data.frame`, you can just call the function `validate()` later on.
Marking observations as validation works analogously to changing column roles:

```{r}
str(task$row_roles)

task$nrow
task$set_row_role(rows = 29:32, new_role = "validation")
task$nrow
```

All pre- and post-processing you have used on the training data is also applied to the validation data in exactly the same way.
